Pottery decor as networks on the Middle Niger

### 2024-04-30 analysis/paper/paper.Rmd
cleaned repo, knit works

#getting absolute numbers of decorations
decors_abs <- decors |> 
  left_join(n_decors) |> 
    mutate(across(where(is.numeric), ~ .*n/100)) |> 
  mutate(across(where(is.numeric), ~ round(.))) |> 
    column_to_rownames("sites") |> 
  select(-n) |> 
t()

iNEXT(decors_abs, q=0, datatype="abundance")

```{r}
 #install_github('AnneChao/iNEXT.4steps') #not currently on CRAN
 library(iNEXT)
 library(iNEXT)
# library(iNEXT.4steps)
 library(ggplot2)

 i <- decors_abs[,c(2,5,8)]|> 
 iNEXT(q=c(0,1,2), datatype="abundance")

 ggiNEXT(i, type=1, facet.var="Assemblage")
ggiNEXT(i, type=3, facet.var="Order.q", color.var="Assemblage")

 i <- decors_abs[,c(2)]|> 
 iNEXT(q=c(0,1,2), datatype="abundance")

 ggiNEXT(i, type=1, facet.var="Assemblage")
ggiNEXT(i, type=3, facet.var="Order.q", color.var="Assemblage")

@@

@@

```{r backbone-eigenvalue, fig.cap="Binary backbone at alpha = 0.15 showing eigenvector centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

@@

```{r backbone-betweenness, fig.cap="Binary backbone at alpha = 0.15 showing betweenness centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_betweenness())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "betweenness centrality")
```

@@

# calculate modularity of louvain clustering
modularity <- decor_graph |>
  activate(nodes) |>
  mutate(modularity = graph_modularity(group = as.factor(louvain))) |>
  pull(modularity) |>
  head(1)
```

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```
```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

# Graph of mean edge weight of the different communities
g1 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = weight, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "mean edge weight",
       x = "community")

# Graph of centralization of the different communities
g2 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

>>>>

#getting absolute numbers of decorations for the 12th century sites
decors_abs <- decors |> 
  left_join(n_decors) |> 
    mutate(across(where(is.numeric), ~ .*n/100)) |> 
  mutate(across(where(is.numeric), ~ round(.))) |> 
    column_to_rownames("sites") |> 
  select(-n) |> 
t()

# getting a table of the results of the analysis
i <- iNEXT(decors_abs, q=0, datatype="abundance")

i

# Fractured and extrapolated diversity of Doupwil
  i <- decors_abs[,c(4)]|> 
 iNEXT(q=c(0), datatype="abundance")
 ggiNEXT(i, type=1, facet.var="Assemblage")+
  labs(title = "Doupwil")

# Fractured and extrapolated diversity of Jenne-jeno
 i <- decors_abs[,c(8)]|> 
 iNEXT(q=c(0), datatype="abundance")
  ggiNEXT(i, type=1, facet.var="Assemblage")+
  labs(title = "Jenne-jeno")

#getting absolute numbers of decorations for the 9th century sites
decors_abs <- decorsearly |> 
  left_join(n_decors) |> 
    mutate(across(where(is.numeric), ~ .*n/100)) |> 
  mutate(across(where(is.numeric), ~ round(.))) |> 
    column_to_rownames("sites") |> 
  select(-n) |> 
t()

# getting a table of the results of the analysis
iNEXT(decors_abs, q=0, datatype="abundance")
```

@@

```
```{r weighted flat decor}
flat_edges <- decor_flat |> 
  get.data.frame("edges")

decor_20 <- decor_20 |> 
  activate(edges) |> 
  select(-weight) |> 
  left_join(flat_edges)

ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
  geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

@@

```{r backbone-eigenvalue, fig.cap="Binary backbone at alpha = 0.15 showing eigenvector centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
   geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

@@

```{r backbone-betweenness, fig.cap="Binary backbone at alpha = 0.15 showing betweenness centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_betweenness())) +
   geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "betweenness centrality")
```

@@

# calculate modularity of louvain clustering
modularity <- decor_20 |>
  activate(nodes) |>
  mutate(modularity = graph_modularity(group = group_louvain())) |>
  pull(modularity) |>
  head(1)
```

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  filter(label != "Sanga D") |> 
  activate(edges) |> 
  #filter(weight > 3) |> 
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.5, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
 flatten(weighted = TRUE) |> 
 activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
flatten(weighted = TRUE) |> 
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.5, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(x = lon, y = lat) +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```
```{r cluster4, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

# Graph of mean edge weight of the different communities
g1 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = weight, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "edge weight",
       x = "community")

# Graph of centralization of the different communities
g2 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "edge div",
       x = "community")
